### 2021-01-03 control-analysis-table.tex
Initial Overleaf Import

\begin{tabular}{lrrrrr}
\toprule{}                    & \textbf{ GLM, 600 vertices} & GLM, 300 vertices & GLM, 900 vertices & Interpolation & GLM, 4 Layers \\
\midrule
Stimulus                      &  {\color[rgb]{0,0.5,0}\textbf{1.648436e-12}} &  {\color[rgb]{0,0.5,0}5.954187e-13} &  {\color[rgb]{0,0.5,0}1.290311e-11} &       {\color[rgb]{0,0.5,0}3.726032e-12} &      {\color[rgb]{0,0.5,0}2.496749e-13} \\
Attention                     &  {\color[rgb]{0,0.5,0}\textbf{1.032463e-08}} &  {\color[rgb]{0,0.5,0}3.833341e-08} &  {\color[rgb]{0,0.5,0}4.175371e-09} &       {\color[rgb]{0,0.5,0}9.482100e-09} &      {\color[rgb]{0,0.5,0}9.216548e-09} \\
Layer                         &  {\color[rgb]{0,0.5,0}\textbf{1.067460e-13}} &  {\color[rgb]{0,0.5,0}2.851381e-11} &  {\color[rgb]{0,0.5,0}7.013292e-13} &       {\color[rgb]{0,0.5,0}1.890989e-15} &      {\color[rgb]{0,0.5,0}3.237563e-15} \\
Area                          &  {\color[rgb]{0,0.5,0}\textbf{1.134083e-08}} &  {\color[rgb]{0,0.5,0}2.712885e-08} &  {\color[rgb]{0,0.5,0}1.114299e-07} &       {\color[rgb]{0,0.5,0}1.019932e-08} &      {\color[rgb]{0,0.5,0}4.101310e-09} \\
Hemisphere                    &  \textbf{9.914955e-01} &  9.461482e-01 &  9.101537e-01 &       9.177039e-01 &      9.029095e-01 \\
Stimulus:Attention            &  \textbf{7.331790e-02} &  2.710442e-01 &  {\color[rgb]{0.5,0,0}2.720432e-02} &       5.523979e-02 &     {\color[rgb]{0.5,0,0}1.179676e-02} \\
Stimulus:Layer                &  {\color[rgb]{0,0.5,0}\textbf{1.433143e-13}} &  {\color[rgb]{0,0.5,0}6.233730e-12} &  {\color[rgb]{0,0.5,0}7.235421e-14} &       {\color[rgb]{0,0.5,0}8.002511e-15} &      {\color[rgb]{0,0.5,0}2.812629e-18} \\
Attention:Layer               &  {\color[rgb]{0,0.5,0}\textbf{8.335165e-10}} &  {\color[rgb]{0,0.5,0}3.172532e-08} &  {\color[rgb]{0,0.5,0}1.934224e-10} &       {\color[rgb]{0,0.5,0}7.100456e-12} &      {\color[rgb]{0,0.5,0}2.061512e-11} \\
Stimulus:Area                 &  {\color[rgb]{0,0.5,0}\textbf{9.424679e-10}} &  {\color[rgb]{0,0.5,0}2.786488e-08} &  {\color[rgb]{0,0.5,0}7.976153e-09} &       {\color[rgb]{0,0.5,0}7.472970e-10} &      {\color[rgb]{0,0.5,0}1.449285e-10} \\
Attention:Area                &  \textbf{7.253885e-02} &  1.208362e-01 &  7.744268e-02 &       1.406810e-01 &      8.801709e-02 \\
Layer:Area                    &  {\color[rgb]{0,0.5,0}\textbf{8.137692e-04}} &  {\color[rgb]{0,0.5,0}3.749919e-03} &  {\color[rgb]{0,0.5,0}7.251511e-03} &       {\color[rgb]{0,0.5,0}1.577094e-05} &      9.033174e-02 \\
Stimulus:Attention:Layer      &  \textbf{3.936910e-01} &  7.659098e-01 &  1.507279e-01 &       7.908435e-02 &      5.661013e-01 \\
Stimulus:Attention:Area       &  \textbf{7.370581e-01} &  6.246065e-01 &  7.617405e-01 &       7.980477e-01 &      7.956297e-01 \\
Stimulus:Layer:Area           &  {\color[rgb]{0.5,0,0}\textbf{2.136465e-02}} &  2.055745e-01 &  {\color[rgb]{0,0.5,0}8.894002e-03} &       {\color[rgb]{0,0.5,0}2.779890e-05} &      {\color[rgb]{0,0.5,0}6.190268e-04} \\
Attention:Layer:Area          &  \textbf{4.155126e-01} &  2.893327e-01 &  3.384936e-01 &       2.462276e-01 &      3.289884e-01 \\
Stimulus:Attention:Layer:Area &  \textbf{8.124587e-01} &  8.742156e-01 &  7.084470e-01 &       9.209281e-01 &      1.570871e-01 \\
\bottomrule
\end{tabular}